---
title: "Using fmaskcheckR"
author: "Bart Huntley"
date: "10 August 2016"
output: html_document
---

## Introduction
The function fmaskcheckR is a helper function designed to iterate through each imagery folder and check if an individual site location falls within a cloud or cloud shadow classified pixel from a previously created fmask cloud mask (see fmaskR doco). If the site is considered cloud free then the image folder name is recorded in a list for that particular site. The function collates these folder names for each site and stores them in a list for input in the jpegR function.

This function was written to be an included option in the jpegR function which produces jpegs of scene dates per site for visual QA for cloud.

This function has been included in the "processingFunctions.R" source script  but can also stand alone as it outputs the collated list to an RData file named "cloudDateSiteList.RData". This RData file can be imported using the load() command and manipulated for other purposes if required.

## Technical - What happens?
When run this function iterates through each scene date in a path/row and:

1. Creates a QAshapes folder in a working directory.

2. Ensures leap year folders are correctly named before proceeding.

3. Splits shape files into single site shapes and stores them in QAshapes folder.

4. Intersects each shape file with each image date cloud mask (output from fmaskR) to determine if completely cloud/shadow free. Handles reprojections if necessary.

5. Builds a list of cloud free image date folders to pass on to jpegR function.

6. Outputs a copy as an RData file to working directory.

## Load fmaskcheckR
Run the below to load:
```{r eval=FALSE}
fmaskcheckR <- function(wkdir, imdir, shp, shp.ID){
  ## Create QAshapes working folder for shapefiles #############################
  setwd(wkdir)
  if(!file.exists("QAshapes")){ dir.create("QAshapes")}
  shpdir <- paste0(wkdir, "\\", "QAshapes")
  
  
  ## libraries #################################################################
  library(raster)
  library(rgdal)
  library(maptools)
  
  ## QA leap file folders for correct date #####################################
  leapR <- function(imdir){
    setwd(imdir)
    allfiles <- list.files(recursive = TRUE)
    result <- allfiles[grepl("*pre.ers", allfiles)]
    result <- result[!grepl("^[a-zA-Z]", result)]
    #date for folder
    fold <- substr(result, 1, 8)
    fdate <- as.character(as.Date(fold, "%Y%m%d"))
    #date for image
    ldate <- as.character(as.Date(substr(result, 21, 26), "%d%m%y"))
    #find mismatch
    bad.fold.dates <- setdiff(fdate, ldate)
    #correct the folder names and path
    corr.fold.dates <- as.Date(bad.fold.dates, "%Y-%m-%d") - 1
    corr.fold <- gsub("-", "", as.character(corr.fold.dates))
    new.name <- paste0(imdir, "\\", corr.fold)
    #old folder names and path to correct
    bad.fold <- gsub("-", "", as.character(bad.fold.dates))
    old.name <- paste0(imdir, "\\", bad.fold)
    #rename folders
    file.rename(old.name, new.name)
  }
  
  leapR(imdir)
  
  
  ## Handle shape file to single parts (sites) #################################
  shpSplitR <- function(wkdir, shpdir, shp, shp.ID){
    setwd(wkdir)
    data <-  readOGR(dsn = ".", shp)
    unique <- unique(data@data[,shp.ID])
    sites <- as.character(unique)
    setwd(shpdir)
    for(i in 1:length(sites)){
      tmp <- data[data@data[,shp.ID] == sites[i], ]
      writeOGR(tmp, dsn=getwd(), sites[i], driver="ESRI Shapefile",
               overwrite_layer=TRUE)
    }
  }
  
  shpSplitR(wkdir, shpdir, shp, shp.ID)
  
  
  ## Get shp names to iterate through ##########################################
  shpfiles <- list.files(path = shpdir, pattern = "*.shp")
  shpfiles <- shpfiles[!grepl("xml", shpfiles)] #xml handler
  shpnames <- unlist(strsplit(shpfiles, split = "\\."))
  shpnames <- shpnames[c(TRUE,FALSE)] #this returns the odd indexes
  
  
  ## Get full list of folders and filepaths ####################################
  list.dirs <- function(path=".", pattern=NULL, all.dirs=FALSE,
                        full.names=FALSE, ignore.case=FALSE) {
    # use full.names=TRUE to pass to file.info
    all <- list.files(path, pattern, all.dirs,
                      full.names=TRUE, recursive=FALSE, ignore.case)
    dirs <- all[file.info(all)$isdir]
    # determine whether to return full names or just dir names
    if(isTRUE(full.names))
      return(dirs)
    else
      return(basename(dirs))
  }
  
  allimdirFolds <- list.dirs(imdir)
  allimdirFolds <- as.Date(allimdirFolds, "%Y%m%d")
  readyFolds <- allimdirFolds[!is.na(allimdirFolds)]
  readyFolds <- gsub("-", "", readyFolds)
  readyFoldsFP <- paste0(imdir, "\\", readyFolds)
  
  
  ## Create list to hold cloud free dates per site #############################
  cloudDateSiteList <- vector("list", length(shpnames))
  goodnames <- paste0("site_", shpnames, "_dates")
  names(cloudDateSiteList) <- goodnames
  
  
  ## Loop through fmask cloud rasters to QA ####################################
  for(i in 1:length(shpnames)){
    siteshp.i <- readOGR(dsn = shpdir, shpnames[i])
    means.i <- numeric(length(readyFoldsFP))
    for(k in 1:length(readyFoldsFP)){
      setwd(readyFoldsFP[k])
      cloud.k <- list.files(pattern = "*cloud.img")
      cloud.k <- cloud.k[!grepl("xml", cloud.k)] #xml handler
      imgcloud.k <- raster(cloud.k)#, band=1
      e.k <- extract(imgcloud.k, siteshp.i)
      means.i[k] <- unlist(lapply(e.k, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
    }
    means.i[means.i != 1] <- 0 
    ind.i <- means.i == 1
    #date.k <- as.Date(readyFolds, "%Y%m%d")
    #date.out <- date.k[ind.i]
    folds.out <- readyFolds[ind.i]
    folds.out <- folds.out[!is.na(folds.out)]
    
    cloudDateSiteList[[i]] <- folds.out
  }
  
  setwd(wkdir)
  save(cloudDateSiteList, file = "cloudDateSiteList.RData")
  return(cloudDateSiteList)
}
```


## Set up parameters

#### wkdir
This is the location of your shape file that contains all the individual sites for analysis. An example might be (note double backslash):
```{r eval=FALSE}
wkdir = "Z:\\project_blah_blah\\working\\analysis"
```

#### imdir
This is the location of the USGS Landsat scene path/row that you wish to process. An example might be (note double backslash):
```{r eval=FALSE}
imdir = "W:\\usgs\\112082"
```

#### shp
This is the name of your shape file that contains all the sites and is located in the working directory. For example (**note no file extension**):
```{r eval=FALSE}
shp = "BartsSites"
```

#### shp.ID
This is the name of the attribute column in your shape file that contains all the sites unique IDs. For example:
```{r eval=FALSE}
shp.ID = "unique_id"
```


## Running fmaskcheckR
Simple matter of:
```{r eval=FALSE}
fmaskcheckR(wkdir, imdir, shp, shp.ID)
```

















